---
title: "ANDREW_DELIS_able.com_case"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
execute:
  warning: false
  message: false
output: html_document
date: "2024-08-09"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

toy_sales_data <- read.csv( "https://raw.githubusercontent.com/jefftwebb/data/main/toy_sales_data.csv" )
toy_sales_PO <- read.csv( "https://raw.githubusercontent.com/jefftwebb/data/main/toy_sales_PO.csv" )

```

### What is the true ATE of having a sale?

```{r }

trueATE <- mean( toy_sales_PO$y1 - toy_sales_PO$y0 )

trueATE

```

### What is the estimated ATE of having a sale using the observed data, toy_sales.csv?

```{r }

treatment_data <- toy_sales_data[toy_sales_data$is_on_sale == 1,]
avg_treatment_sales <- mean( treatment_data$weekly_amount_sold )

control_data <- toy_sales_data[toy_sales_data$is_on_sale == 0,]
avg_control_sales <- mean( control_data$weekly_amount_sold )

estimatedATE <- avg_treatment_sales - avg_control_sales

estimatedATE

```

### What is the bias involved in using the observed data to estimate ATE?

```{r }

bias <- estimatedATE - trueATE

bias

```

### A balance table using the observed data (toy_sales.csv) that shows how company size varies by treatment status. Comment on the exchangeability of the groups

```{r }

bt_data <- toy_sales_data |>
  group_by( is_on_sale ) |>
  summarize( avg_co_size = mean( avg_week_sales ) )

# Shows that the average company size varies per group 
# This is evidence that these groups are not perfectly exchangable. 
bt_data$avg_co_size[2] - bt_data$avg_co_size[1]

bt_data

```

### Some analysis to show how company size is affecting the relationship between weekly_amount_sold and is_on_sale, and a statistical model to adjust for the confounding.  

```{r }

lm( weekly_amount_sold ~ is_on_sale + avg_week_sales, data = toy_sales_data ) |>
  summary()

```

#### Adjusted ATE

```{r }

# Create a new column in a new data frame
adjusted_data <- mutate( toy_sales_data, adjusted_sales = weekly_amount_sold - ( avg_week_sales * 3.1902 ) )

# Find the ATE using adjusted sales column
treatment_data_2 <- adjusted_data[adjusted_data$is_on_sale == 1,]
avg_treatment_sales_2 <- mean( treatment_data_2$adjusted_sales )

control_data_2 <- adjusted_data[adjusted_data$is_on_sale == 0,]
avg_control_sales_2 <- mean( control_data_2$adjusted_sales )

estimatedATE_2 <- avg_treatment_sales_2 - avg_control_sales_2

estimatedATE_2

```

#### How the adjustment is working.

Based on the linear regression, sales increases 3.1902 for every unit increase in company size (average weekly sales). To adjust for that I created a new column called "adjusted_sales" to the observed data that removed that effect by multiplying the average weekly sales by the linear regression coefficient given and then subtracted that from the total sales per unit. Then I found the ATE again.

### What is the true ATE of a sales campaign conditional on weeks to Christmas? 

```{r }

CATE_data <- toy_sales_PO |>
  group_by( weeks_to_xmas ) |>
  summarize( CATE = mean( y1 - y0 ) )

CATE_data

```

### A statistical model used to estimate the true ATE of a sales campaign conditional on weeks to Christmas.

```{r }

# Run regression model
lm( weekly_amount_sold ~ is_on_sale * weeks_to_xmas, data = toy_sales_data ) |>
  summary()

# Create a new column in a new data frame
interaction_data <- mutate( toy_sales_data, adjusted_sales = weekly_amount_sold - ( weeks_to_xmas * 6.942 ) )

# Find the ATE using adjusted sales column 

treatment_data_3 <- interaction_data[interaction_data$is_on_sale == 1,]
avg_treatment_sales_3 <- mean( treatment_data_3$adjusted_sales )

control_data_3 <- interaction_data[interaction_data$is_on_sale == 0,]
avg_control_sales_3 <- mean( control_data_3$adjusted_sales )

estimatedATE_3 <- avg_treatment_sales_3 - avg_control_sales_3

estimatedATE_3
```
